\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-12 podc16/paper2.tex
added a liveness property to map-agreement

\begin{enumerate}[noitemsep,nolistsep]
    \item[-] For any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
    \item[-] For any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.

\end{enumerate}

@@

\begin{comment}
Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
\begin{itemize}[noitemsep,nolistsep]
    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
\end{itemize}

The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
Its key property is \cref{lemma1}. 
We assume that processes initially agree, for each set of commands $\mathcal{D}$, on a total order $<_{\mathcal{D}}$ on $\mathcal{D}$.
In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{D}}$ can be the restriction of this total order on $\mathcal{D}$.

%\textbf{Local dependency graph computed by a process.} 
We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$). For example, if $m_p = \left[ c_1 \mapsto \aset{c_2,c_4}, c_2 \mapsto \aset{c_3}, c_3 \mapsto \aset{c_2}\right]$ then $V\left( g_p \right) = \aset{c_2,c_3}$ and $E\left( g_p \right) = \aset{\left( c_2,c_3 \right),\left( c_3,c_2 \right)}$.

%\textbf{Orders induced by a graph.}
A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and none from $c_2$ to $c_1$.
%Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
%If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
%\textbf{Linearizations of a dependency graph.}
We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{D}$ and $ c_1 <_{\mathcal{D}} c_2$ hold, then $c_1 \ll c_2$. For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
%\textbf{Partial order on graphs.}

@@

\begin{comment}

\textbf{Graph of strongly connected components.} Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ (represented as sets of vertices) and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$. Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order. 
For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. Then $V(\hat{h}) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E( \hat{h} )=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$. We will use $h$ as an example in the rest of the section.

\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ is a total order on the set $\mathcal{D}$.

\textbf{Linearizations of a dependency graph.}
When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, we say that a total order $l$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when $\left(v_1,v_2\right) \in l$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
$v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
For example, if $f\left[ \aset{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $h$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$

\begin{comment}
Define $\opn{Lin}_g\left( \leq \right) = f\left[ V_1 \right] \cdot F\left[ V_2 \right] \cdot F\left[ V_n \right]$, where $V_i\in V\left( \hat{g} \right)$

Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ as the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ... \todo{or first define the partial order induced by a graph and then its compatible total orders.}

Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ contains each command in $\mathcal{D}$ exactly once.
A sequence of commands $cseq$ is an $f$-linearization of the graph $g$ if and only there is a sequence whose element of strongly connected

\todo{Explain the graph processing algo independently? Its key property is that when $G1 \subseteq G2$ then $lin\left( c_1,G1 \right)$ is a prefix of $lin\left( c_2,G2 \right)$ or vice-versa.}

@@

\textbf{Lemma 2}
If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.

@@

The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons. Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced. Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}). Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.

>>>>

\begin{enumerate}[noitemsep,nolistsep]
    \item[-] Safety: (a), for any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (b), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.  
    \item[-] Liveness: if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
\end{enumerate}

@@

The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
Its key property is \cref{lemma1}. 
We assume that processes initially agree, for each set of commands $\mathcal{D}$, on a total order $<_{\mathcal{D}}$ on $\mathcal{D}$.
In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{D}}$ can be the restriction of this total order to $\mathcal{D}$.

We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$). For example, if $m_p = \left[ c_1 \mapsto \aset{c_2,c_4}, c_2 \mapsto \aset{c_3}, c_3 \mapsto \aset{c_2}\right]$ then $V\left( g_p \right) = \aset{c_2,c_3}$ and $E\left( g_p \right) = \aset{\left( c_2,c_3 \right),\left( c_3,c_2 \right)}$.

A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and none from $c_2$ to $c_1$.
For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{D}$ and $ c_1 <_{\mathcal{D}} c_2$ hold, then $c_1 \ll c_2$. For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.

@@

@@

\begin{mylemma}\label{lemma-compat}
If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
\end{mylemma}
\Cref{lemma-compat} follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.

@@

The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons. Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command iff there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced. Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}). Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
